---
title: "BN Notes"
subtitle: Exploratory notes and analysis for Beyond Notability
listing:
  - id: everything
    contents:
    - "notes/*.qmd"
    - posts
    sort: "date desc"
    exclude: 
      title: "Other Notes"
    type: table
    fields: [date, title, file-modified, categories]
    field-display-names:
      file-modified: "Updated"
    categories: true
    sort-ui: false
    filter-ui: false
    date-format: "D MMM YYYY"
    max-items: 99
toc: false
page-layout: full
title-block-banner: false
execute: 
  warning: false
  message: false
  echo: false 
---

The site does not contain live wikibase data but will be quite frequently updated. (Doing the updates is straightforward but as the number of pages grows it takes longer to run a full update, so it'll probably happen less often. Let me know if you want to see an updated version of a page at any time.) 

```{r}
# will need to check this actually works...
#latest_knit <- Sys.Date()
#Latest data update: ``r latest_knit``

## ok consolidating notes and posts sseems to work, but keep original separate listings code below for reference

## want to show everything (it's not likely to get silly and that should fix filtering problems) set max-items at 99, unlikely to end up with more than that...!

## original listings

  # - id: other-notes
  #   contents: "notes/*.qmd"
  #   sort: "title"
  #   exclude: 
  #     title: "Other Notes"
  #   type: table
  #   fields: [title, file-modified, categories]
  #   field-display-names:
  #     file-modified: "Updated"
  #   categories: false
  #   sort-ui: false
  #   filter-ui: false
  #   date-format: "D MMM YYYY"
  # - id: posts-table
  #   contents: posts
  #   sort: "date desc"
  # 
  #   type: table
  #   fields: [date, title, file-modified, categories]
  #   field-display-names:
  #     file-modified: "Updated"
  #   categories: true
  #   sort-ui: false
  #   filter-ui: false
  #   date-format: "D MMM YYYY"

```


Caveats: 

This is all work in progress, provisional, generally handle with caution, etc. But, because the wikibase itself is continually evolving, updates may sometimes break things without me noticing, so if you see anything that really looks *strange*, let me know.

Known issues:

As of December 2023, "The Bug" (recent edits *are* in the wikibase but for some reason the query service fails to index them) has been fixed. This means, in particular, dates of birth and death should now be all up to date. The problem may well recur in the future and I'll be checking from time to time. 


## Posts and Notes

[Posts](blog.qmd): short pieces on specific topics or questions 

[Notes](notes/): overviews, summary tables and other Stuff

(The Updated column in this listing refers to modifications to the file, not data updates. The Modified field on individual pages should be the most recent data update for that page.)

:::{#everything}
:::


```{r separate-listings-deprecated}
# ## Notes
# 
# Overviews, summary tables and other documents 
# 
# :::{#other-notes}
# :::
# 
# [All notes](notes/)
# 
# 
# ## Recent Posts
# 
# Short posts on specific questions or topics
# 
# ::: {#posts-table}
# :::
# 
# [All posts](blog.qmd)

```



