---
title: "PPAs and change over time"
description: "Swarms!"
author: "Sharon Howard"
date: "2023-12-12"
date-modified: today
categories: 
  - ppa
  - dates
execute: 
  warning: false
  message: false
  echo: false 
---

## Wikibase queries

* [Dates](https://beyond-notability.wikibase.cloud/wiki/Project:SPARQL/examples#Dates) section
* [PPA buckets](https://beyond-notability.wikibase.cloud/wiki/Project:SPARQL/examples#PPA_Buckets_for_Women)

```{r}
# this doesn't have a Table
# find out what's happened to date_prop_label. (was only used in reactable i think, but still)
```


```{r libraries-etc}
# shared libraries, functions etc ####

source(here::here("_R/shared.R"))

# any extra libraries will usually go here 

library(ggbeeswarm)
library(ggridges)

# standard query strings and queries ####

source(here::here("_R/std_queries.R"))

```





```{r dates-data}
source(here::here("_R/dates.R"))


bn_women_dates_birth <-
bn_women_dates |>
  filter(date_prop %in% c("P26")) |>
  #add_count(bn_id) |> filter(n>1) |>
    group_by(bn_id) |>
  arrange(date_precision, .by_group = TRUE) |>
  top_n(1, row_number()) |>
  ungroup()

bn_women_dates_categories_birth <-
  bn_women_dates_categories |>
  # might as well add dob here...? will need a filter for dob specific charts
  left_join(
    bn_women_dates_birth |>
      select(bn_id, date_birth=date, year_birth=year), by="bn_id"
  ) |>
  mutate(age = year - year_birth) |>
  mutate(period = case_when(
    between(year_birth, 1815, 1854) ~ "1815-1853",
    between(year_birth, 1855, 1877) ~ "1854-1876",
    between(year_birth, 1878, 1924) ~ "1877-1924"
  )) 
# earlier grouping. added a lot of dates since then!
  # mutate(period = case_when(
  #   between(year_birth, 1825, 1858) ~ "1825-1858",
  #   between(year_birth, 1855, 1881) ~ "1859-1881",
  #   between(year_birth, 1881, 1924) ~ "1882-1924"
  # ))


bn_women_dates_ppa <-
bn_women_dates_categories_birth |>
  filter(category %in% c("PPA")) |>
  # this should be all you need now...
  # need shorter labels for event participation and charitable/organisational !
  mutate(subcat = case_match(
    ppa_bucket,
    "event participation" ~ "events",
    "charitable/organisational" ~ "charitable",
    .default = ppa_bucket
  )) 


```





```{r}
## but... how many of these have more than one dates essentially for the same event, ie start and end? 
## hmm it might not matter too much unless you have very disproportionate coverage of pit v s/e in different categories?
```




## Notes


>Can you make a bee swarm-type diagram that plots the frequency of PPAs across all the women in the dataset, across the entire project period? So with years ascending up the y axis, and all the women considered one big blotchy group (am sure that's the technical term). What we want to get is an overview of whether certain kinds of PPA were more dominant at various periods of the overall period covered by the project, e.g. women suddenly started doing more excavation or whatnot in the 1910s, whereas before that there was almost no excavation activity associated with women. Project data on PPAs still very incomplete for post-1920s period, at least until CAS Earthworks and Blue Paper data is fully added, but even if we could begin to get a sense for the 1870 - 1920 period this would be very helpful to us. 


- Updated to use the PPA buckets
- I've moved the Big Blob versions to the bottom of the page since there was a clear preference for the separated approach!


```{r}
#This uses a slightly different layout method that forces all the swarms to the same width (so big categories are denser but don't overlap their neighbours which was happening with spoke at).

# bn_women_dates_ppa |>
#   filter(between(year, 1871, 1950))  |>
#   filter(!subcat %in% c("membership", "other", "resources")) |>
# #  filter(!date_prop %in% c("P75", "P67") & subcat!="other") |>
#   arrange(year, subcat) |>
#   ggplot(aes(y=year, x=subcat)) +
#   ggbeeswarm::geom_quasirandom(size=0.4,  varwidth = TRUE) +  #, corral = "wrap", corral.width = 0.9, cex=0.8
#   #scale_colour_ptol() +
#   #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
#   theme(legend.position = "bottom") +
#   guides(colour = guide_legend(override.aes = list(size=2))) +
#   labs(y=NULL, x=NULL, colour=NULL)

```

## Overview

The PPA buckets ([see also](./bn_notes/posts/ppa-2024-02-25/)).

```{r}
# most compact approach? try a table with bucket in LH column and categories horizontally in RH ?
bn_ppa_buckets |>
  filter(!label %in% c("no buckets")) |>
  group_by(label) |>
  arrange(str_to_lower(ppa), .by_group = T) |>
  summarise(ppa = paste(ppa, collapse = " ~ ") ) |>
  ungroup() |>
  rename(bucket=label, activity=ppa) |>
  kable() |>
  kable_styling()
```



## Beeswarms

### dates

```{r}

bn_women_dates_ppa |>
  filter(between(year, 1861, 1950))  |> 
  filter(!subcat %in% c("charitable", "no buckets")) |>
#  filter(!date_prop %in% c("P75", "P67") & subcat!="other") |>
  arrange(year, subcat) |>
  ggplot(aes(y=year, x=subcat)) +
  ggbeeswarm::geom_beeswarm(size=0.4,  method = "swarm", cex = 0.8, corral = "wrap", corral.width = 0.9) + #, cex=0.8
  scale_colour_ptol() +
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(size=2))) +
  labs(y=NULL, x=NULL, colour=NULL)

```

Although it becomes (I think) less effective with very small categories, we can drill down into different properties in a larger category. 

in the case of positions, most are `held position`

```{r}

bn_women_dates_ppa |>
  filter(between(year, 1861, 1950))  |>
  filter(subcat =="positions") |>
#  filter(!date_prop %in% c("P75", "P67") & subcat!="other") |>
  #arrange(year, subcat) |>
  ggplot(aes(y=year, x=date_prop_label_std)) +
  ggbeeswarm::geom_beeswarm(size=0.5,  method = "swarm", cex = 1.2) +  #, corral = "wrap", corral.width = 0.9, cex=0.8
  scale_colour_ptol() +
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(size=2))) +
  labs(y=NULL, x=NULL, colour=NULL, title="positions")

```


```{r}

bn_women_dates_ppa |>
  filter(between(year, 1861, 1950))  |>
  filter(subcat =="elections") |>
#  filter(!date_prop %in% c("P75", "P67") & subcat!="other") |>
  #arrange(year, subcat) |>
  ggplot(aes(y=year, x=date_prop_label_std)) +
  ggbeeswarm::geom_beeswarm(size=0.5,  method = "swarm", cex = 1.2) +  #, corral = "wrap", corral.width = 0.9, cex=0.8
  scale_colour_ptol() +
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(size=2))) +
  labs(y=NULL, x=NULL, colour=NULL, title = "elections")

```

### age


```{r}
bn_women_dates_ppa |>
  filter(between(year, 1861, 1950) & !is.na(age))  |>
  filter(!subcat %in% c("charitable", "no buckets")) |>
#  filter(!date_prop %in% c("P75", "P67") & subcat!="other") |>
  arrange(age, subcat) |>
  ggplot(aes(y=age, x=subcat)) +
  ggbeeswarm::geom_beeswarm(size=0.4,  method = "swarm", cex = 0.8, corral = "wrap", corral.width = 0.9) + #, cex=0.8
  #scale_colour_ptol() +
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  theme(legend.position = "bottom") +
  #guides(colour = guide_legend(override.aes = list(size=2))) +
  labs(y=NULL, x=NULL, colour=NULL)
```

### age and birth period

This splits the women into age groups; I've reintroduced a colour element because the facets are too narrow for the category labels.

```{r}
bn_women_dates_ppa |> 
  filter(between(year, 1861, 1950) & !is.na(age))  |> 
  filter(!is.na(period)) |>
  filter(!subcat %in% c("charitable", "no buckets")) |>
  arrange(age, subcat) |>
  ggplot(aes(y=age, x=subcat, colour=subcat)) +
  ggbeeswarm::geom_beeswarm(size=0.2,  method = "swarm", cex = 0.8, corral = "wrap", corral.width = 0.9) + #, cex=0.8
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  scale_colour_ptol() +
  facet_wrap(~period) +
  guides(colour = guide_legend(override.aes = list(size=2))) +
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_blank()) +
  labs(y=NULL, x=NULL, colour=NULL)
```

```{r}
# bn_women_dates_ppa |> 
#   filter(between(year, 1861, 1950) & !is.na(age))  |> 
#   filter(!is.na(period)) |>
#   filter(!subcat %in% c("charitable", "no buckets")) |>
#   distinct(bn_id, period) |>
#   count(period)
# 1815-1853	152			
# 1854-1876	148			
# 1877-1924	136
```




```{r fig.height=7.5}
# bn_women_dates_ppa_education_dob |>
#   filter(between(year, 1871, 1950))  |> 
#   filter(!is.na(period)) |>
#   filter(!subcat %in% c("membership", "other", "resources", "education")) |>
# #  filter(!date_prop %in% c("P75", "P67") & subcat!="other") |>
#   arrange(age, subcat) |>
#   mutate(subcat = fct_rev(subcat)) |>
#   ggplot(aes(x=age, y=subcat, colour=subcat)) +
#   ggbeeswarm::geom_beeswarm(size=0.2,  method = "swarm", cex = 0.8, corral = "wrap", corral.width = 0.9, show.legend = F) + #, cex=0.8
#   #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
#   scale_colour_ptol() +
#   facet_wrap(~period, ncol=1) +
#   #guides(colour = guide_legend(override.aes = list(size=2))) +
#   theme(#legend.position = "bottom"
#     aspect.ratio = 1
#         ) +
#   #theme(axis.text.x = element_blank()) +
#   labs(y=NULL, x=NULL, colour=NULL)
```


### playing around with variables


```{r}

bn_women_dates_ppa |>
  filter(between(year, 1861, 1950) )  |> 
  filter(!is.na(period)) |>
  filter(!subcat %in% c("charitable", "no buckets")) |>
  arrange(year, subcat) |>
  ggplot(aes(y=year, x=subcat, colour=subcat)) +
  ggbeeswarm::geom_beeswarm(size=0.4,  method = "swarm", cex = 0.8, corral = "wrap", corral.width = 0.9) + #, cex=0.8
  facet_wrap(~period) +
  scale_colour_ptol() +
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  theme(axis.text.x = element_blank()) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(size=2))) +
  labs(y=NULL, x=NULL, colour=NULL)

#ggsave("test-ppa-bees.png", width=7.3, height=4.5)
```

```{r}

bn_women_dates_ppa |>
  filter(between(year, 1861, 1950) )  |> 
  filter(!is.na(period)) |>
  filter(!subcat %in% c("charitable", "no buckets")) |>
  arrange(age, subcat) |>
  ggplot(aes(y=age, x=period, colour=period)) +
  ggbeeswarm::geom_beeswarm(size=0.4,  method = "swarm", cex = 0.8, corral = "wrap", corral.width = 0.9) + #, cex=0.8
  facet_wrap(~subcat, ncol=4) +
  scale_colour_colorblind() +
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  theme(axis.text.x = element_blank()) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(size=2))) +
  labs(y=NULL, x=NULL, colour=NULL)

#ggsave("test-ppa-bees.png", width=7.3, height=4.5)
```

### held position

This only shows job titles with at least 10 dates. (I've included employed as for the handful that met the threshold.)

```{r}
# # why not do this with employed as well? then colour 
# bn_women_dates_ppa_education |>
#   filter(str_detect(date_prop_label_std, "held position") & !str_detect(prop_valueLabel, "job title unrecorded|^t\\d")) |>
#   select(bn_id, personLabel, prop_valueLabel, date_prop_label, date_label, year, s) |>
#   add_count(prop_valueLabel, date_prop_label, sort = T) |>
#   filter(n>=10) |>
#   ggplot(aes(x=year, y=prop_valueLabel)) +
#   ggbeeswarm::geom_beeswarm(size=0.4,  method = "swarm", cex = 0.8, corral = "wrap", corral.width = 0.9) + #, cex=0.8
#   #scale_colour_ptol() +
#   #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
#   theme(legend.position = "bottom") +
#   #guides(colour = guide_legend(override.aes = list(size=2))) +
#   labs(y=NULL, x=NULL, colour=NULL)
```

```{r}
bn_women_dates_ppa |>
  filter(str_detect(date_prop_label_std, "held position|employed as") & !str_detect(prop_valueLabel, "job title unrecorded|^t\\d")) |>
  select(bn_id, personLabel, prop_valueLabel, #date_prop_label, 
         date_prop_label_std, date_label, year, s) |>
  add_count(prop_valueLabel, date_prop_label_std, sort = T) |>
  filter(n>=10) |>
  mutate(prop_valueLabel = fct_rev(fct_relevel(prop_valueLabel, "Honorary Secretary", "Local Secretary", "Secretary", "Academic", "Lecturer", "Lecturer (freelance)", "Lecturer (Extension)"))) |>
  ggplot(aes(x=year, y=prop_valueLabel, colour=date_prop_label_std)) +
  ggbeeswarm::geom_beeswarm(size=0.4,  method = "swarm", cex = 0.8, corral = "wrap", corral.width = 0.9) + #, cex=0.8
  scale_colour_tableau() +
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(size=1))) +
  labs(y=NULL, x=NULL, colour=NULL, caption = "at least 10 dates")  

# maybe a consolidated version?
# ok maybe leave for now, but possibly want to do some reordering to group things together.
# assistant (x) keep separate from (x) ? 
# what about honorary (x) ? there are a lot of honorary secretary but few in other jobs
```

And by age, in birth groups. 

```{r}
bn_women_dates_ppa |>
  filter(str_detect(date_prop_label_std, "held position|employed as") & !str_detect(prop_valueLabel, "job title unrecorded|^t\\d")) |>
  select(bn_id, personLabel, age, period, prop_valueLabel, #date_prop_label, 
         date_prop_label_std, date_label, year, s) |>
  add_count(prop_valueLabel, date_prop_label_std, sort = T) |>
  filter(!is.na(period)) |>
  filter(n>=10) |>
  mutate(prop_valueLabel = fct_rev(fct_relevel(prop_valueLabel, "Honorary Secretary", "Local Secretary", "Secretary", "Academic", "Lecturer", "Lecturer (freelance)", "Lecturer (Extension)"))) |>
  ggplot(aes(x=age, y=prop_valueLabel, colour=date_prop_label_std)) +
  ggbeeswarm::geom_beeswarm(size=0.4,  method = "swarm", cex = 0.8, corral = "wrap", corral.width = 0.9) + #, cex=0.8
  scale_colour_tableau() +
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  facet_wrap(~period) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(size=1))) +
  labs(y=NULL, x=NULL, colour=NULL, caption = "at least 10 dates")  
```


```{r}
# Assistant
# County Collector
# Assistant Curator
# Curator
# Chairman
# Director
# Editor
# Honorary Editor
# Honorary Treasurer
# Inspector
# Academic
# Lecturer
# Lecturer (Extension)
# Lecturer (freelance)
# Teacher
# Tutor
# Archivist
# Cataloguer
# Librarian
# Assistant Secretary
# Honorary Secretary
# Local Secretary
# Secretary
# President
# Vice President

```



## Ridgeline plots

These are another option for this kind of comparison of trends. They're also not dependent on colour, so it's not necessary to reduce categories quite so brutally. I've included education this time. 

They show the *proportion* of dates (rather than counts) in each "bin" (5?-year-period) for each category. So categories are directly comparable regardless of size (though very small ones can be a bit misleading).

I've experimented with using colour as an indication of the size of the category - the lighter the shade of blue the larger the category (but I've done it in a really hacky sort of way and would need to fix it properly).

```{r}
#  adapted from smallbills commodities
bn_women_dates_ppa |>
  filter(between(year, 1861, 1950) )  |> 
  filter(!is.na(period)) |>
  filter(!subcat %in% c("charitable", "no buckets")) |>
  add_count(subcat) |>
 mutate(subcat = fct_rev(subcat)) |>
  # how to use fill like a heatmap? stat(x) in one example but doesn't work here
 ggplot(aes(x = year, y = subcat, fill=n, group=subcat)) +   ##, height=after_stat(density))) +  
  geom_density_ridges(stat = "binline", bins = 18, scale = 0.95, draw_baseline = FALSE, alpha=0.7) +
  # will generally have to set the `expand` option for both axes to remove unneeded padding 
  scale_y_discrete(expand = c(0, 0)) +     
  scale_x_continuous(expand = c(0, 0), breaks = seq(1870, 1950, 20) ) +  
  #scale_fill_viridis_b() +
  coord_cartesian(clip = "off") +  #clip = "off" to avoid clipping of the very top of the top ridgeline
  theme_ridges(font_size=12) +
  theme(axis.text.y = element_text(vjust = 0), legend.position = "none", 
        panel.grid.minor.x=element_blank(),
        panel.grid.major.x=element_blank()
        ) +
  labs(x=NULL, y=NULL)
```

Showing individual properties (with 50+ instances) (some light categorisation might still be useful though, I think)
  
```{r}
#  adapted from smallbills commodities
bn_women_dates_ppa |>
  filter(between(year, 1861, 1950) )  |> 
  filter(!is.na(period)) |>
  #filter(!subcat %in% c("charitable", "no buckets")) |>
  add_count(date_prop_label_std, subcat, sort = T) |>
  filter(n>50) |>
 mutate(date_prop_label_std = fct_rev(date_prop_label_std)) |>
  # how to use fill like a heatmap? stat(x) in one example but doesn't work here
 ggplot(aes(x = year, y = date_prop_label_std, fill=n, group=date_prop_label_std)) +   ##, height=after_stat(density))) +  
  geom_density_ridges(stat = "binline", bins = 18, scale = 0.95, draw_baseline = FALSE, alpha=0.7) +
  # will generally have to set the `expand` option for both axes to remove unneeded padding 
  scale_y_discrete(expand = c(0, 0)) +     
  scale_x_continuous(expand = c(0, 0), breaks = seq(1870, 1950, 20) ) +   
  coord_cartesian(clip = "off") +  #clip = "off" to avoid clipping of the very top of the top ridgeline
  theme_ridges(font_size=12) +
  theme(axis.text.y = element_text(vjust = 0), legend.position = "none", 
        panel.grid.minor.x=element_blank(),
        panel.grid.major.x=element_blank()
        ) +
  labs(x=NULL, y=NULL)
```


This is a variation on the same theme. The ridgeline shows a smoothed trend line, but this adds the actual distribution of dates as points, so it's a bit more like the beeswarm. Colours are just for fun this time around.

```{r}

bn_women_dates_ppa |>
  filter(between(year, 1861, 1950) )  |> 
  filter(!is.na(period)) |>
  filter(!subcat %in% c("charitable", "no buckets")) |>
 mutate(subcat = fct_rev(subcat)) |>
 ggplot(aes(x = year, y = subcat, fill=subcat, group=subcat)) +   ##, height=after_stat(density))) +  
  geom_density_ridges(jittered_points = TRUE, position = "points_sina",scale = 0.95, alpha=0.6, point_size=.5, bandwidth=4) +
  coord_cartesian(clip = "off") +  #clip = "off" to avoid clipping of the very top of the top ridgeline
  # will generally have to set the `expand` option for both axes to remove unneeded padding 
  scale_y_discrete(expand = c(0, 0)) +     
  scale_x_continuous(expand = c(0, 0), breaks = seq(1870, 1950, 20) ) +   
  theme_ridges(font_size=12) +
  theme(axis.text.y = element_text(vjust = 0), legend.position = "none", 
        panel.grid.minor.x=element_blank(),
        panel.grid.major.x=element_blank()
        ) +
  labs(x=NULL, y=NULL)
```


## The Blob

- I've flipped the axes for this because it makes better use of space
- It's quite pretty but I'm not convinced it's the most effective way to visualise what you want to know 
  - the focus of the earlier beeswarms was comparison of different groups rather than looking closely within a single blob
  - there are so many PPA categories vs how many colours it's possible to use in one chart

```{r}

bn_women_dates_ppa |>
  filter(between(year, 1861, 1950) )  |> 
  #filter(!is.na(period)) |>
  filter(!subcat %in% c("charitable", "no buckets")) |>
  arrange(year, subcat) |>
  ggplot(aes(year, "", colour=subcat)) +
  ggbeeswarm::geom_beeswarm(size=0.5,  method = "swarm", cex = 1.2) +  #, corral = "wrap", corral.width = 0.9, cex=0.8
  scale_colour_ptol() +
  #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(size=2))) +
  labs(y=NULL, x=NULL, colour=NULL)

#ggsave("test-ppa-bees.png", width=7.3, height=4.5)
```


```{r}
##interactive version. this is rubbish. 

# bn_ppa_bees <-
# bn_women_dates_ppa |>
#   filter(between(year, 1861, 1950) )  |> 
#   #filter(!is.na(period)) |>
#   filter(!subcat %in% c("charitable", "no buckets")) |>
#   arrange(year, subcat) |>
#   ggplot(aes(year, "", colour=subcat)) +
#   geom_beeswarm(size=0.5,  method = "swarm", cex = 1.2) +  #, corral = "wrap", corral.width = 0.9, cex=0.8
#   scale_colour_ptol() +
#   #scale_x_discrete(expand=expansion(add=c(0.5,0.5))) +
#   #theme(legend.position = "bottom") +  # plotly ignores this 
#   #guides(colour = guide_legend(override.aes = list(size=2))) + # ?conflict with plotly? legend doesn't show at all.
#   labs(y=NULL, x=NULL)
# 
# plotly::ggplotly(bn_ppa_bees)
```


## Table


```{r}
# what's happened to date_prop_label ??
bn_women_dates_reactable <-
bn_women_dates_ppa |>
  select(bn_id, personLabel, ppa_bucket, date_prop_label_std, prop_valueLabel, year, date_prop, person) |>
  arrange(bn_id, year, ppa_bucket, date_prop)
```

```{r}

reactable(
  bn_women_dates_reactable |> select(-person, -date_prop) ,
  filterable = TRUE,
  searchable = TRUE,
  bordered = TRUE,
  highlight = TRUE,
  defaultPageSize = 130,
  #groupBy = "event",
  defaultExpanded = TRUE,
  defaultColDef = colDef(maxWidth = 60),  # , headerClass = "sticky" but i think it has to be fixed height to use this
  columns = list(
    bn_id = colDef(name="id", maxWidth = 70),
    personLabel =colDef(maxWidth = 200, name="name",
      html = TRUE,
      cell = function(value, index) {
        sprintf('<a href="%s" target="_blank">%s</a>', bn_women_dates_reactable$person[index], value)
      })   ,
    ppa_bucket = colDef(maxWidth = 140, name = "PPA bucket"),
    date_prop_label_std = colDef(maxWidth = 160, name="activity"),
    prop_valueLabel = colDef(maxWidth = 180, name="what/where"),
    # collegeLabel = colDef(maxWidth = 250, name="college",
    #   html = TRUE,
    #   cell = function(value, index) {
    #     sprintf('<a href="%s" target="_blank">%s</a>', bn_fsa_education$college[index], value)
    #   })   ,  # TODO work out how to make this conditional
   year=colDef(minWidth = 90)
    
  )
)
```

