---
title: "Work activities (P17/P48/P105)"
description: "A look at work activities with the properties `held position` and `employed as`"
author: "Sharon Howard"
date: "2023-08-22"
date-modified: today
categories: 
  - work
  - ppa
execute: 
  warning: false
  message: false
  echo: false 
page-layout: full
---


```{r libraries-etc}
# shared libraries, functions etc ####

source(here::here("_R/shared.R"))

# any extra libraries will usually go here 

# standard query strings and queries ####

source(here::here("_R/std_queries.R"))

```


```{r new-query}
bn_women_work_activities_new_query <-
c(paste(
  bn_prefixes,
        '
SELECT distinct ?personLabel ?work_label ?positionLabel ?date_label ?date 
?employerLabel ?organisedLabel ?ofLabel 
?s ?person ?work ?position ?employer ?organised ?of 

WHERE {
    
  ?person bnwdt:P3 bnwd:Q3 . # women
  
  # get work activities: held position / held position (free text) /  employed as
  ?person ( bnp:P17|bnp:P48|bnp:P105 ) ?s .  
  
  ?person ?work_p ?s . # for activity type label
  ?work wikibase:claim ?work_p;       
         rdfs:label ?work_label. filter(lang(?work_label)="en-gb") .  
    
    # more about the position
    ?s ( bnps:P17|bnps:P48|bnps:P105 ) ?position .  
    # employer / organised by / of (incl free text). TODO: check whether these are mutually exclusive; could make this a union.
    OPTIONAL { ?s bnpq:P18 ?employer .}  
    OPTIONAL { ?s bnpq:P109 ?organised .}
    OPTIONAL { ?s ( bnpq:P78 | bnpq:P66 ) ?of .}
 
   # optional { ?s bnpq:P2 ?location . } # sometimes there is location and no employer/org/of 
    
  # dates
    optional {
    ?s ?pq ?date . 
      
      ?qual_prop wikibase:qualifier ?pq;      
              #wikibase:propertyType ?date_prop_type; # can check date property types if needed
              wikibase:propertyType wikibase:Time ; # nb *does not* include edtf dates but AFAICT there are none in qualifiers.
              rdfs:label ?date_label . filter(lang(?date_label)="en-gb") . # what kind of date is it.
      } # /dates
   
  SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en-gb,en". } 
  
} # /where

ORDER BY ?person
')) |>
  sparql2df(endpoint=bn_endpoint) |>
  make_bn_item_id(person)  |>
  # make blanks NA
  mutate(across(c(-bn_id, -personLabel), ~na_if(., "")))  

bn_women_work_activities <-
bn_women_work_activities_new_query |>
  # make <unknown value> NA - strings start with _:t
  mutate(date = if_else(str_detect(date, "^_:t"), NA, date))|>
  mutate(date = parse_date_time(date, "ymdHMS")) |>
  mutate(year = year(date)) |>
  mutate(s_id = str_remove(s, "https://beyond-notability.wikibase.cloud/entity/statement/"))  |>
  # drop employer/org/of for now, a few dups and not needed for this. 
  distinct(bn_id, personLabel, work_label, positionLabel, date_label, date, year, s_id, work, person, position) |>
  # keep NAs for now but drop end time
  filter(date_label !="end time" | is.na(date_label))  |>
  mutate(employed_position = case_when(
    work_label=="employed as" ~ work_label,
    str_detect(work_label, "held position") ~ "held position"
  ))
# still a handful with multi dates for same s, but i thnk can be ignored.
# note. number of employed as is unchanged since first iteration but there are 414 position cf originally 398.
```




## Wikibase query

[Work](https://beyond-notability.wikibase.cloud/wiki/Project:SPARQL/examples#Work) - [Women's work activities with employers and dates](https://beyond-notability.wikibase.cloud/wiki/Project:SPARQL/examples#Women_and_their_work_activities,_with_employers/organisations_and_dates)

Note: Updated 12/9/23 to use a new version of the example queries that captures more date types.

## Next steps...

- employers/organisations/locations
- wider range of public and professional activities
- requests????

## Chronological

[Using date or start date, whichever is present. 52/372 `position held` and 14/75 `employed as` have no date information.]

The big spike in `position held` is in 1892. Why?


```{r}
bn_women_work_activities |>
  count(year, employed_position) |>
  filter(!is.na(year)) |>
  ggplot(aes(year, n)) +
  geom_col() +
  facet_wrap(~employed_position, scales = "free_y") +
  labs(title="Jobs: position held and employed as, by year", y="count of jobs", x="year")
```




Percentage of posts that are `employed as` (with trend line)

```{r}

# bn_women_employed_positions |>
#   count(year_date1_n, type) |>
#   filter(!is.na(year_date1_n)) |>
#   group_by(year_date1_n) |>
#   mutate(p = n/sum(n)) |>
#   ungroup() |>
#   complete(year_date1_n, type,
#            fill=list(p=0)) |>
#   filter(type=="employed as") |>
#   ggplot(aes(year_date1_n, p)) +
#   geom_point() +
#   geom_smooth() +
#   scale_y_continuous(labels = percent_format()) +
#   labs(title="'Employed as' posts per year", x="year", y="% of jobs")
```


```{r}

bn_women_work_activities |>
  count(year, employed_position)  |>
  filter(!is.na(year)) |>
  group_by(year) |>
  mutate(p = n/sum(n)) |>
  ungroup() |>
  complete(year, employed_position,
           fill=list(p=0)) |>
  filter(employed_position=="employed as") |>
  ggplot(aes(year, p)) +
  geom_point() +
  geom_smooth() +
  scale_y_continuous(labels = percent_format()) +
  labs(title="'Employed as' posts per year", x="year", y="% of jobs")
```





## Job descriptions

[Slight reduction: removed "assistant" and "honorary" qualifiers]

```{r}
# problem with job description of County Collector because it lacks en-gb label and auto_description doens't work in R.
# annoying but unless you can work out how to find the missing en-gb labels, you'll have to make sure you add en to queries here
bn_women_work_activities |>
  #count(year, employed_position) bn_women_employed_positions |>
  mutate(position_label = case_when(
    str_detect(positionLabel, "^t\\d+") ~ "na",
    positionLabel=="job title unrecorded" ~ "na",
    .default = str_to_lower(positionLabel)
  ))  |>
  mutate(position_label = str_trim(str_remove(position_label, "^(assistant|honorary) +"))) |>
  count(position_label, employed_position) |>
  mutate(t = sum(n), .by = position_label) |>
  filter(position_label !="na", t>1) |>
  mutate(position_label = fct_reorder(position_label, t)) |>
  ggplot() +
  geom_col(aes(x=position_label, y=n, fill=employed_position)) +
  coord_flip() +
  scale_fill_tableau() +
  labs(x=NULL, y=NULL, title="Job descriptions, ordered by frequency", fill="type", caption = "count>1")
```






## Individual women



```{r}
bn_women_work_activities |>
  count(personLabel, employed_position)  |>
  mutate(t = sum(n), .by = personLabel) |>
  mutate(personLabel = fct_reorder(personLabel, t)) |>
  filter(t>3) |>
  ggplot() +
  geom_col(aes(x=personLabel, y=n, fill=employed_position)) +
  coord_flip() +
  scale_fill_tableau() +
  labs(x=NULL, y=NULL, title="Individual women's job counts, ordered by frequency", fill="type", caption = "count>3")
```





## Table


```{r}
# need to check this, the unnest may cause problems.
## wtf is wrong with pivot wider? surely it doesn't normall do this...
# need end dates again oops. don't need amalgamated work type though
bn_women_work_activities_reactable <-
bn_women_work_activities_new_query |>
  # make <unknown value> NA - strings start with _:t
  mutate(date = if_else(str_detect(date, "^_:t"), NA, date))|>
  mutate(date = parse_date_time(date, "ymdHMS")) |>
  mutate(year = year(date)) |>
  mutate(s_id = str_remove(s, "https://beyond-notability.wikibase.cloud/entity/statement/"))  |>
  # changes
  # keep only what you need for the table 
  distinct(bn_id, personLabel, work_label, positionLabel, date_label, date, s_id, person) |>
# mark the undated rows
  mutate(nd = if_else(is.na(date), "no date", NA)) |>
  # fkg pivot_wider
  pivot_wider(names_from = date_label, values_from = date, values_fn = list) |>
  clean_names("snake") |>
  # unnest the list cols pivot wider introduced; pretty sure this works
  unnest(point_in_time:latest_date, keep_empty = TRUE) |>
  select(-s_id, -na) 
  
# you can't use arrange inside the reactable, you idiot
bn_women_work_activities_reactable_sorted <-
bn_women_work_activities_reactable |>
    mutate(date_sort = case_when(
      !is.na(point_in_time) ~ point_in_time,
      !is.na(start_time) ~ start_time,
      !is.na(latest_date) ~ latest_date
    )) |>
  arrange( parse_number(str_remove(bn_id, "^Q") ), date_sort) 
```





```{r}

reactable(
  bn_women_work_activities_reactable_sorted|>
  select(-person, -bn_id, -date_sort)   ,
  filterable = TRUE,
  searchable = TRUE,
  bordered = TRUE,
  highlight = TRUE,
  defaultPageSize = 100,
  #groupBy = "event",
  defaultExpanded = TRUE,
  columns = list(
    #bn_id = colDef(name="id", maxWidth = 70),
    person_label =colDef(maxWidth = 150, name="name",
      html = TRUE,
      cell = function(value, index) {
        sprintf('<a href="%s" target="_blank">%s</a>', bn_women_work_activities_reactable_sorted$person[index], value)
      })   ,
    point_in_time = colDef(minWidth = 75, name="date", format = colFormat(date = TRUE)),
    start_time = colDef(minWidth = 75, name="start", format = colFormat(date = TRUE)),
    end_time = colDef(minWidth = 75, name="end", format = colFormat(date = TRUE)),
    latest_date = colDef(name="latest", minWidth = 75, format = colFormat(date=TRUE)),
    work_label = colDef(maxWidth = 100, name="type"),
    position_label = colDef(maxWidth = 180, name="post"),
    nd = colDef(maxWidth = 60)
    #employerLabel = colDef(maxWidth = 130, name="employer"),
  )
)
```


